---
title: "BTC CytAssist Pilot"
author: "LTKAGOHARA"
date: "2024-02-26"
output:
  html_document: default
  pdf_document: default
---

# VISIUM DATA ANALYSIS WITH SEURAT
# Sample BTC01 (pilot)

##### Setting working directory
```{r}
setwd("~/OneDrive - Johns Hopkins/Visium_BTC_DPT/BTC_Dec2023/")
```

##### Loading packages
```{r}
library(Seurat)
library(ggplot2)
library(leiden)
library(patchwork)
library(dplyr)
library(rhdf5)
library(ggrepel)
library(Rfast2)
library(EnhancedVolcano)
library(fgsea)
library(msigdbr)
library(RColorBrewer)
library(ggpubr)
library(ComplexHeatmap)
library(DESeq2)
```

## Increasing the allowed memory to avoid errors regarding lack of memory
```{r}
set.seed(1234)
options(future.globals.maxSize = 1024 * 1024 * 1024)  # 1 GB
```

## Loading the data for sample BTC01
```{r}
set.seed(1234)
BTC01 <- Load10X_Spatial("BTC_DPTpilot1_spaceranger/BTC01/")
dim(BTC01)
head(BTC01@meta.data)
```

## INTEGRATING CODA CELL CLASSIFICATION INTO METADATA
```{r}
set.seed(1234)

# extract BTC01 metadata
BTC01metadata <- BTC01@meta.data

# uploading CODA annotations
codaBTC01 <- as.data.frame(read.csv("Analysis_pilot/BTC01/tissue_positions_cellular_compositions.csv"))
rownames(codaBTC01) <- codaBTC01$X
codaBTC01 <- codaBTC01[, -c(1, 2, 3, 4, 5, 6)]
head(codaBTC01)

# merging CODA annotations ot metadata
BTC01metadata <- merge(BTC01metadata, codaBTC01, by = "row.names")

BTC01metadata$CODA <- "mixed"
BTC01metadata$CODA[BTC01metadata$RespondingPDAC > 51] <- "responding"
BTC01metadata$CODA[BTC01metadata$NonrespondingPDAC > 51] <- "non-responding"
BTC01metadata$CODA[BTC01metadata$Stroma > 51] <- "stroma"
BTC01metadata$CODA[BTC01metadata$Immune > 51] <- "immune"
BTC01metadata$CODA[BTC01metadata$Vasculature > 51] <- "vasculature"
BTC01metadata$CODA[BTC01metadata$Fat > 51] <- "fat"
BTC01metadata$CODA[BTC01metadata$Acini > 51] <- "acini"
BTC01metadata$CODA[BTC01metadata$Islets > 51] <- "islets"
BTC01metadata$CODA[BTC01metadata$Nerve > 51] <- "nerve"
BTC01metadata$CODA[BTC01metadata$Duct > 51] <- "normal duct"
BTC01metadata$CODA[BTC01metadata$RespondingPDAC == 0 & 
                 BTC01metadata$NonrespondingPDAC == 0 & 
                 BTC01metadata$Stroma == 0 & 
                 BTC01metadata$Immune == 0 &
                 BTC01metadata$Vasculature == 0 &
                 BTC01metadata$Fat == 0 &
                 BTC01metadata$Acini == 0 &
                 BTC01metadata$Islets == 0 &
                 BTC01metadata$Nerve == 0 &
                 BTC01metadata$Duct == 0] <- "NA"
table(BTC01metadata$CODA)

BTC01 <- AddMetaData(object = BTC01, metadata = BTC01metadata)
head(BTC01)
```

##### saving object with CODA classification
```{r}
save(BTC01, file = "Analysis_pilot/BTC01/BTC01_CODA.Rda")
```

# DATA PREPROCESSING

## QC and filtering
```{r}
set.seed(1234)

load("Analysis_pilot/BTC01/BTC01_CODA.Rda")

# proportion of mitochondrial reads
BTC01[["percent_mt"]] <- PercentageFeatureSet(BTC01, pattern = "^MT-")
head(BTC01@meta.data)

VlnPlot(BTC01, features = c("nFeature_Spatial", "nCount_Spatial", "percent_mt"), ncol = 3)
ggsave("Analysis_pilot/BTC01/BTC01_QC1.png")

plot1 <- FeatureScatter(BTC01, feature1 = "nCount_Spatial", feature2 = "percent_mt")
plot2 <- FeatureScatter(BTC01, feature1 = "nCount_Spatial", feature2 = "nFeature_Spatial")
plot1 | plot2
ggsave("Analysis_pilot/BTC01/BTC01_QC2.png")

# plots to verify the variance in counts across spots
plot3 <- VlnPlot(BTC01, features = "nCount_Spatial", pt.size = 0.1) + NoLegend()
plot4 <- SpatialFeaturePlot(BTC01, features = "nCount_Spatial", pt.size.factor = 7) + theme(legend.position = "right")
plot3 | plot4
ggsave("Analysis_pilot/BTC01/BTC01_QC3.png")

# NOTE: data looks good quality and spots were not filtered out of the analysis
```

##### saving filtered dataset
```{r}
save(BTC01, file = "Analysis_pilot/BTC01/BTC01_filtered.Rda")
```

## NORMALIZATION WITH SCTransform
```{r}
set.seed(1234)

load("Analysis_pilot/BTC01/BTC01_filtered.Rda")

BTC01 <- SCTransform(BTC01, assay = "Spatial", vars.to.regress = "percent_mt",verbose = F)
```

##### saving SCTransformed data
```{r}
save(BTC01, file = "Analysis_pilot/BTC01/BTC01_SCTransformed.Rda")
```

## Loading the data for sample BTC03
```{r}
set.seed(1234)
BTC03 <- Load10X_Spatial("BTC_DPTpilot1_spaceranger/BTC03/")
dim(BTC03)
head(BTC03@meta.data)
```

## INTEGRATING CODA CELL CLASSIFICATION INTO METADATA
```{r}
set.seed(1234)

# extract BTC03 metadata
BTC03metadata <- BTC03@meta.data

# uploading CODA annotations
codaBTC03 <- as.data.frame(read.csv("Analysis_pilot/BTC03/tissue_positions_cellular_compositions.csv"))
rownames(codaBTC03) <- codaBTC03$X
codaBTC03 <- codaBTC03[, -c(1, 2, 3, 4, 5, 6)]
head(codaBTC03)

# merging CODA annotations ot metadata
BTC03metadata <- merge(BTC03metadata, codaBTC03, by = "row.names")

BTC03metadata$CODA <- "mixed"
BTC03metadata$CODA[BTC03metadata$RespondingPDAC > 51] <- "responding"
BTC03metadata$CODA[BTC03metadata$NonrespondingPDAC > 51] <- "non-responding"
BTC03metadata$CODA[BTC03metadata$Stroma > 51] <- "stroma"
BTC03metadata$CODA[BTC03metadata$Immune > 51] <- "immune"
BTC03metadata$CODA[BTC03metadata$Vasculature > 51] <- "vasculature"
BTC03metadata$CODA[BTC03metadata$Fat > 51] <- "fat"
BTC03metadata$CODA[BTC03metadata$Acini > 51] <- "acini"
BTC03metadata$CODA[BTC03metadata$Islets > 51] <- "islets"
BTC03metadata$CODA[BTC03metadata$Nerve > 51] <- "nerve"
BTC03metadata$CODA[BTC03metadata$Duct > 51] <- "normal duct"
BTC03metadata$CODA[BTC03metadata$RespondingPDAC == 0 & 
                 BTC03metadata$NonrespondingPDAC == 0 & 
                 BTC03metadata$Stroma == 0 & 
                 BTC03metadata$Immune == 0 &
                 BTC03metadata$Vasculature == 0 &
                 BTC03metadata$Fat == 0 &
                 BTC03metadata$Acini == 0 &
                 BTC03metadata$Islets == 0 &
                 BTC03metadata$Nerve == 0 &
                 BTC03metadata$Duct == 0] <- "NA"
table(BTC03metadata$CODA)

BTC03 <- AddMetaData(object = BTC03, metadata = BTC03metadata)
head(BTC03)
```

##### saving object with CODA classification
```{r}
save(BTC03, file = "Analysis_pilot/BTC03/BTC03_CODA.Rda")
```

# DATA PREPROCESSING

## QC and filtering
```{r}
set.seed(1234)

load("Analysis_pilot/BTC03/BTC03_CODA.Rda")

# proportion of mitochondrial reads
BTC03[["percent_mt"]] <- PercentageFeatureSet(BTC03, pattern = "^MT-")
head(BTC03@meta.data)

VlnPlot(BTC03, features = c("nFeature_Spatial", "nCount_Spatial", "percent_mt"), ncol = 3)
ggsave("Analysis_pilot/BTC03/BTC03_QC1.png")

plot1 <- FeatureScatter(BTC03, feature1 = "nCount_Spatial", feature2 = "percent_mt")
plot2 <- FeatureScatter(BTC03, feature1 = "nCount_Spatial", feature2 = "nFeature_Spatial")
plot1 | plot2
ggsave("Analysis_pilot/BTC03/BTC03_QC2.png")

# plots to verify the variance in counts across spots
plot3 <- VlnPlot(BTC03, features = "nCount_Spatial", pt.size = 0.1) + NoLegend()
plot4 <- SpatialFeaturePlot(BTC03, features = "nCount_Spatial", pt.size.factor = 7) + theme(legend.position = "right")
plot3 | plot4
ggsave("Analysis_pilot/BTC03/BTC03_QC3.png")

# NOTE: data looks good quality and spots were not filtered out of the analysis
```

##### saving filtered dataset
```{r}
save(BTC03, file = "Analysis_pilot/BTC03/BTC03_filtered.Rda")
```

## NORMALIZATION WITH SCTransform
```{r}
set.seed(1234)

load("Analysis_pilot/BTC03/BTC03_filtered.Rda")

BTC03 <- SCTransform(BTC03, assay = "Spatial", vars.to.regress = "percent_mt",verbose = F)
```

##### saving SCTransformed data
```{r}
save(BTC03, file = "Analysis_pilot/BTC03/BTC03_SCTransformed.Rda")
```

## Loading the data for sample BTC09
```{r}
set.seed(1234)
BTC09 <- Load10X_Spatial("BTC_DPTpilot1_spaceranger/BTC09/")
dim(BTC09)
head(BTC09@meta.data)
```

## INTEGRATING CODA CELL CLASSIFICATION INTO METADATA
```{r}
set.seed(1234)

# extract BTC09 metadata
BTC09metadata <- BTC09@meta.data

# uploading CODA annotations
codaBTC09 <- as.data.frame(read.csv("Analysis_pilot/BTC09/tissue_positions_cellular_compositions.csv"))
rownames(codaBTC09) <- codaBTC09$X
codaBTC09 <- codaBTC09[, -c(1, 2, 3, 4, 5, 6)]
head(codaBTC09)

# merging CODA annotations ot metadata
BTC09metadata <- merge(BTC09metadata, codaBTC09, by = "row.names")

BTC09metadata$CODA <- "mixed"
BTC09metadata$CODA[BTC09metadata$RespondingPDAC > 51] <- "responding"
BTC09metadata$CODA[BTC09metadata$NonrespondingPDAC > 51] <- "non-responding"
BTC09metadata$CODA[BTC09metadata$Stroma > 51] <- "stroma"
BTC09metadata$CODA[BTC09metadata$Immune > 51] <- "immune"
BTC09metadata$CODA[BTC09metadata$Vasculature > 51] <- "vasculature"
BTC09metadata$CODA[BTC09metadata$Fat > 51] <- "fat"
BTC09metadata$CODA[BTC09metadata$Acini > 51] <- "acini"
BTC09metadata$CODA[BTC09metadata$Islets > 51] <- "islets"
BTC09metadata$CODA[BTC09metadata$Nerve > 51] <- "nerve"
BTC09metadata$CODA[BTC09metadata$Duct > 51] <- "normal duct"
BTC09metadata$CODA[BTC09metadata$RespondingPDAC == 0 & 
                 BTC09metadata$NonrespondingPDAC == 0 & 
                 BTC09metadata$Stroma == 0 & 
                 BTC09metadata$Immune == 0 &
                 BTC09metadata$Vasculature == 0 &
                 BTC09metadata$Fat == 0 &
                 BTC09metadata$Acini == 0 &
                 BTC09metadata$Islets == 0 &
                 BTC09metadata$Nerve == 0 &
                 BTC09metadata$Duct == 0] <- "NA"
table(BTC09metadata$CODA)

BTC09 <- AddMetaData(object = BTC09, metadata = BTC09metadata)
head(BTC09)
```

##### saving object with CODA classification
```{r}
save(BTC09, file = "Analysis_pilot/BTC09/BTC09_CODA.Rda")
```

# DATA PREPROCESSING

## QC and filtering
```{r}
set.seed(1234)

load("Analysis_pilot/BTC09/BTC09_CODA.Rda")

# proportion of mitochondrial reads
BTC09[["percent_mt"]] <- PercentageFeatureSet(BTC09, pattern = "^MT-")
head(BTC09@meta.data)

VlnPlot(BTC09, features = c("nFeature_Spatial", "nCount_Spatial", "percent_mt"), ncol = 3)
ggsave("Analysis_pilot/BTC09/BTC09_QC1.png")

plot1 <- FeatureScatter(BTC09, feature1 = "nCount_Spatial", feature2 = "percent_mt")
plot2 <- FeatureScatter(BTC09, feature1 = "nCount_Spatial", feature2 = "nFeature_Spatial")
plot1 | plot2
ggsave("Analysis_pilot/BTC09/BTC09_QC2.png")

# plots to verify the variance in counts across spots
plot3 <- VlnPlot(BTC09, features = "nCount_Spatial", pt.size = 0.1) + NoLegend()
plot4 <- SpatialFeaturePlot(BTC09, features = "nCount_Spatial", pt.size.factor = 7) + theme(legend.position = "right")
plot3 | plot4
ggsave("Analysis_pilot/BTC09/BTC09_QC3.png")

# NOTE: data looks good quality and spots were not filtered out of the analysis
```

##### saving filtered dataset
```{r}
save(BTC09, file = "Analysis_pilot/BTC09/BTC09_filtered.Rda")
```

## NORMALIZATION WITH SCTransform
```{r}
set.seed(1234)

load("Analysis_pilot/BTC09/BTC09_filtered.Rda")

BTC09 <- SCTransform(BTC09, assay = "Spatial", vars.to.regress = "percent_mt",verbose = F)
```

##### saving SCTransformed data
```{r}
save(BTC09, file = "Analysis_pilot/BTC09/BTC09_SCTransformed.Rda")
```

## Loading the data for sample BTC13
```{r}
set.seed(1234)
BTC13 <- Load10X_Spatial("BTC_DPTpilot1_spaceranger/BTC13/")
dim(BTC13)
head(BTC13@meta.data)
```

## INTEGRATING CODA CELL CLASSIFICATION INTO METADATA
```{r}
set.seed(1234)

# extract BTC13 metadata
BTC13metadata <- BTC13@meta.data

# uploading CODA annotations
codaBTC13 <- as.data.frame(read.csv("Analysis_pilot/BTC13/tissue_positions_cellular_compositions.csv"))
rownames(codaBTC13) <- codaBTC13$X
codaBTC13 <- codaBTC13[, -c(1, 2, 3, 4, 5, 6)]
head(codaBTC13)

# merging CODA annotations ot metadata
BTC13metadata <- merge(BTC13metadata, codaBTC13, by = "row.names")

BTC13metadata$CODA <- "mixed"
BTC13metadata$CODA[BTC13metadata$RespondingPDAC > 51] <- "responding"
BTC13metadata$CODA[BTC13metadata$NonrespondingPDAC > 51] <- "non-responding"
BTC13metadata$CODA[BTC13metadata$Stroma > 51] <- "stroma"
BTC13metadata$CODA[BTC13metadata$Immune > 51] <- "immune"
BTC13metadata$CODA[BTC13metadata$Vasculature > 51] <- "vasculature"
BTC13metadata$CODA[BTC13metadata$Fat > 51] <- "fat"
BTC13metadata$CODA[BTC13metadata$Acini > 51] <- "acini"
BTC13metadata$CODA[BTC13metadata$Islets > 51] <- "islets"
BTC13metadata$CODA[BTC13metadata$Nerve > 51] <- "nerve"
BTC13metadata$CODA[BTC13metadata$Duct > 51] <- "normal duct"
BTC13metadata$CODA[BTC13metadata$RespondingPDAC == 0 & 
                 BTC13metadata$NonrespondingPDAC == 0 & 
                 BTC13metadata$Stroma == 0 & 
                 BTC13metadata$Immune == 0 &
                 BTC13metadata$Vasculature == 0 &
                 BTC13metadata$Fat == 0 &
                 BTC13metadata$Acini == 0 &
                 BTC13metadata$Islets == 0 &
                 BTC13metadata$Nerve == 0 &
                 BTC13metadata$Duct == 0] <- "NA"
table(BTC13metadata$CODA)

BTC13 <- AddMetaData(object = BTC13, metadata = BTC13metadata)
head(BTC13)
```

##### saving object with CODA classification
```{r}
save(BTC13, file = "Analysis_pilot/BTC13/BTC13_CODA.Rda")
```

# DATA PREPROCESSING

## QC and filtering
```{r}
set.seed(1234)

load("Analysis_pilot/BTC13/BTC13_CODA.Rda")

# proportion of mitochondrial reads
BTC13[["percent_mt"]] <- PercentageFeatureSet(BTC13, pattern = "^MT-")
head(BTC13@meta.data)

VlnPlot(BTC13, features = c("nFeature_Spatial", "nCount_Spatial", "percent_mt"), ncol = 3)
ggsave("Analysis_pilot/BTC13/BTC13_QC1.png")

plot1 <- FeatureScatter(BTC13, feature1 = "nCount_Spatial", feature2 = "percent_mt")
plot2 <- FeatureScatter(BTC13, feature1 = "nCount_Spatial", feature2 = "nFeature_Spatial")
plot1 | plot2
ggsave("Analysis_pilot/BTC13/BTC13_QC2.png")

# plots to verify the variance in counts across spots
plot3 <- VlnPlot(BTC13, features = "nCount_Spatial", pt.size = 0.1) + NoLegend()
plot4 <- SpatialFeaturePlot(BTC13, features = "nCount_Spatial", pt.size.factor = 7) + theme(legend.position = "right")
plot3 | plot4
ggsave("Analysis_pilot/BTC13/BTC13_QC3.png")

# NOTE: data looks good quality and spots were not filtered out of the analysis
```

##### saving filtered dataset
```{r}
save(BTC13, file = "Analysis_pilot/BTC13/BTC13_filtered.Rda")
```

## NORMALIZATION WITH SCTransform
```{r}
set.seed(1234)

load("Analysis_pilot/BTC13/BTC13_filtered.Rda")

BTC13 <- SCTransform(BTC13, assay = "Spatial", vars.to.regress = "percent_mt",verbose = F)
```

##### saving SCTransformed data
```{r}
save(BTC13, file = "Analysis_pilot/BTC13/BTC13_SCTransformed.Rda")
```

```{r}
sessionInfo()
```
