---
title: "BTC DPT Visium HD Cohort - SDC Classification to EMT"
author: "Meredith Wetzel"
date: "2025-10-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = "15_figures/",
  dev = "pdf")
```

```{r message=FALSE, warning=FALSE}
rm(list = ls())
library(Seurat)
library(ggplot2)
library(patchwork)
library(dplyr)
library(RColorBrewer)
library(EnhancedVolcano)
library(fgsea)
library(DESeq2)
library(DT)
library(ggbiplot)
library(reshape2)
library(ggpubr)
library(CoGAPS)
library(SeuratDisk)
source(file = "../../../stacked_props_function.R")
options(future.globals.maxSize = 8000 * 1024^2)
set.seed(1234)
```

```{r include=FALSE}
obj.dir <- "/projects/btcpilot_visiumhd/objects/"
```

```{r}
# load(file = paste0(obj.dir,"noHC2_merged_016um_cirroRCTD_seurObj_updated_VBannotCombined.rda"))
load(file = paste0(obj.dir,"final_codeCheck/allSubtypes_fullObj_016um_seurObj.rda"))

# names(sobj@images) <- unique(sobj$sample)
sobj <- JoinLayers(sobj, assay = "Spatial.016um", layers = "data")
```

```{r}
rctd.cols <- c('red','blue','green2','pink','yellow','orange','turquoise','purple','gold2','lightblue','lightgreen','magenta','tan','lightgrey')
names(rctd.cols) <- c("ENDOTHELIAL","FIBROBLASTS","PDAC","MYELOID","PLASMA","PERICYTES","CYCLING MYELOID","ENDOCRINE","TNK","ACINAR","B CELLS","MAST","CYCLING TNK","NA")
celltypes <- c("ENDOTHELIAL","FIBROBLASTS","PDAC","MYELOID","PLASMA","PERICYTES","CYCLING MYELOID","ENDOCRINE","TNK","ACINAR","B CELLS","MAST","CYCLING TNK","NA")

sub.cols <- c("green4","blue2","maroon1","darkturquoise","mediumpurple1","orange")
names(sub.cols) <- c("Basal","Classical","myCAF","iCAF","Intermediate","CAF")

caf.type.color <- c("maroon1", "orange", "darkturquoise")
names(caf.type.color) <- c("myCAF", "CAF", "iCAF")

epi.type.color <- c("green4","blue2","mediumpurple1")
names(epi.type.color) <- c("Basal","Classical","Intermediate")

response.cols <- c(rgb(53,150,54,maxColorValue = 255),rgb(227,172,43,maxColorValue = 255)) # to match the SD colors/Andre's
names(response.cols) <- c("non-responding","responding")

```

# Overall Cells Plot

```{r fig.height=10, fig.width=15, warning=FALSE}
SpatialDimPlot(sobj[,sobj$cell_type %in% celltypes], label = F, ncol = 3, pt.size.factor = 8, image.alpha = 0.5, group.by = "cell_type", cols = rctd.cols)
```

# Classifications - per spot

```{r}
# get SDC1 and SDC4 median (non-zero) cutoffs:
# subset to just PDAC:
Idents(sobj) <- "cell_type"
sobj <- subset(sobj, idents = c("PDAC"))

SDC1_exp <- sobj@assays$Spatial.016um@layers$data[rownames(sobj) == 'SDC1',] # normalized data
sobj$SDC1_exp <- SDC1_exp

SDC4_exp <- sobj@assays$Spatial.016um@layers$data[rownames(sobj) == 'SDC4',]
sobj$SDC4_exp <- SDC4_exp

median(SDC1_exp[SDC1_exp != 0])
median(SDC4_exp[SDC4_exp != 0])

# per spot - using spot median
cutoff <- median(SDC1_exp[SDC1_exp != 0])
print(paste0("Median, non-zero, expression of SDC1 for PDAC cells: ", cutoff))

sobj$SDC1_class <- sapply(sobj$SDC1_exp, function (x){
  if(x >= cutoff){
    return("SDC1_hi")
  }
  if(x < cutoff){
    return("SDC1_lo")
  }
})

cutoff <- median(SDC4_exp[SDC4_exp != 0])
print(paste0("Median, non-zero, expression of SDC4 for PDAC cells: ", cutoff))

sobj$SDC4_class <- sapply(sobj$SDC4_exp, function (x){
  if(x >= cutoff){
    return("SDC4_hi")
  }
  if(x < cutoff){
    return("SDC4_lo")
  }
})

table(sobj$SDC1_class, sobj$CODA)
table(sobj$SDC4_class, sobj$CODA)

table(sobj$SDC1_class, sobj$sample)
table(sobj$SDC4_class, sobj$sample)

sobj$sampleCODA <- paste0(sobj$sample,"_",sobj$CODA)

table(sobj$SDC1_class, sobj$sampleCODA)
table(sobj$SDC4_class, sobj$sampleCODA)

```

```{r fig.height=10, fig.width=15}
class.cols <- c("gold","navy","gold","navy")
names(class.cols) <- c("SDC1_hi","SDC1_lo","SDC4_hi","SDC4_lo")

SpatialDimPlot(sobj, group.by = "SDC1_class", pt.size.factor = 7, ncol = 3, cols = class.cols)
SpatialDimPlot(sobj, group.by = "SDC4_class", pt.size.factor = 7, ncol = 3, cols = class.cols)
```

```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
# EMT module score needs to be added
HALLMARK <- gmtPathways("../../../h.all.v2023.2.Hs.symbols.gmt")
for(i in seq_along(names(HALLMARK))){
  path <- names(HALLMARK[i])
  genes <- HALLMARK[i]
  sobj <- AddModuleScore(sobj, features = genes, name = path)
}
```

# SDC1
```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
comparisons <- list(c("SDC1_hi","SDC1_lo"))
plot_list_SDC1 <- list()
for(i in seq_along(names(HALLMARK))){
  path <- paste0(names(HALLMARK[i]),"1")
  p <- VlnPlot(sobj, features = path, group.by = "SDC1_class", cols = class.cols, pt.size = 0) + geom_boxplot(width=0.1, fill="white") + NoLegend() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),axis.title.x=element_blank(),axis.title.y=element_blank(), legend.title = element_text(size = 8), legend.text = element_text(size = 6), plot.title = element_text(size = 10), axis.text = element_text(size = 8))+ stat_compare_means(method = "t.test", comparisons = comparisons, paired = FALSE) + scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))
  plot_list_SDC1[[path]] <- p
  #print(p)
}# these stats/pvalues are NOT adjusted
plot_list_SDC1[["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1"]]
```

# SDC4
```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
comparisons <- list(c("SDC4_hi","SDC4_lo"))
plot_list_SDC4 <- list()
for(i in seq_along(names(HALLMARK))){
  path <- paste0(names(HALLMARK[i]),"1")
  p <- VlnPlot(sobj, features = path, group.by = "SDC4_class", cols = class.cols, pt.size = 0) + geom_boxplot(width=0.1, fill="white") + NoLegend() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),axis.title.x=element_blank(),axis.title.y=element_blank(), legend.title = element_text(size = 8), legend.text = element_text(size = 6), plot.title = element_text(size = 10), axis.text = element_text(size = 8))+ stat_compare_means(method = "t.test", comparisons = comparisons, paired = FALSE) + scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))
  plot_list_SDC4[[path]] <- p
  #print(p)
}# these stats/pvalues are NOT adjusted
plot_list_SDC4[["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1"]]
```

# split response plots

# SDC1
```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
comparisons <- list(c("SDC1_hi","SDC1_lo"))
plot_list_SDC1 <- list()
for(i in seq_along(names(HALLMARK))){
  path <- paste0(names(HALLMARK[i]),"1")
  p <- VlnPlot(sobj, features = path, group.by = "SDC1_class", split.by = "responseType", pt.size = 0, cols = response.cols) + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),axis.title.x=element_blank(),axis.title.y=element_blank(), legend.title = element_text(size = 8), legend.text = element_text(size = 6), plot.title = element_text(size = 10), axis.text = element_text(size = 8))
  plot_list_SDC1[[path]] <- p
  #print(p)
}# these stats/pvalues are NOT adjusted
plot_list_SDC1[["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1"]]
```

# SDC4
```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
comparisons <- list(c("SDC4_hi","SDC4_lo"))
plot_list_SDC4 <- list()
for(i in seq_along(names(HALLMARK))){
  path <- paste0(names(HALLMARK[i]),"1")
  p <- VlnPlot(sobj, features = path, group.by = "SDC4_class", pt.size = 0, split.by = "responseType", cols = response.cols)+ theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),axis.title.x=element_blank(),axis.title.y=element_blank(), legend.title = element_text(size = 8), legend.text = element_text(size = 6), plot.title = element_text(size = 10), axis.text = element_text(size = 8))
  plot_list_SDC4[[path]] <- p
  #print(p)
}# these stats/pvalues are NOT adjusted
plot_list_SDC4[["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1"]]
```

# Aggregated (BULK) cutoff/classification

```{r}
#aggregate based on sample and CODA annot
bulk_sobj <- AggregateExpression(sobj, assays = "Spatial.016um", return.seurat = T, group.by = c('sample','cell_type'))
info <- bulk_sobj@meta.data
```

```{r}
#create column for SDC1 expression
SDC1_exp_bulk <- bulk_sobj@assays$Spatial.016um$data[rownames(bulk_sobj) == 'SDC1',]
bulk_sobj$SDC1_exp_bulk <- SDC1_exp_bulk

SDC4_exp_bulk <- bulk_sobj@assays$Spatial.016um$data[rownames(bulk_sobj) == 'SDC4',]
bulk_sobj$SDC4_exp_bulk <- SDC4_exp_bulk
```

```{r}
cutoff <- median(SDC1_exp_bulk[SDC1_exp_bulk != 0])
print(paste0("Median, non-zero, expression of SDC1 for PDAC cells in bulk: ", cutoff))
```

```{r}
### ADD CLASSIFICATIONS to Seurat Object using BULK cutoff
sobj$SDC1_class_bulk <- sapply(sobj$SDC1_exp, function (x){
  if(x >= cutoff){
    return("SDC1_hi")
  }
  if(x < cutoff){
    return("SDC1_lo")
  }
})
```

```{r}
cutoff <- median(SDC4_exp_bulk[SDC4_exp_bulk != 0])
print(paste0("Median, non-zero, expression of SDC4 for PDAC cells in bulk: ", cutoff))
```

```{r}
### ADD CLASSIFICATIONS to Seurat Object using BULK cutoff
sobj$SDC4_class_bulk <- sapply(sobj$SDC4_exp, function (x){
  if(x >= cutoff){
    return("SDC4_hi")
  }
  if(x < cutoff){
    return("SDC4_lo")
  }
})
```

```{r fig.height=10, fig.width=15}
SpatialDimPlot(sobj, group.by = "SDC1_class_bulk", pt.size.factor = 7, ncol = 3, cols = class.cols)
SpatialDimPlot(sobj, group.by = "SDC4_class_bulk", pt.size.factor = 7, ncol = 3, cols = class.cols)
```


# SDC1 bulk
```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
comparisons <- list(c("SDC1_hi","SDC1_lo"))
plot_list_SDC1_bulk <- list()
for(i in seq_along(names(HALLMARK))){
  path <- paste0(names(HALLMARK[i]),"1")
  p <- VlnPlot(sobj, features = path, group.by = "SDC1_class_bulk", cols = class.cols, pt.size = 0) + geom_boxplot(width=0.1, fill="white") + NoLegend() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),axis.title.x=element_blank(),axis.title.y=element_blank(), legend.title = element_text(size = 8), legend.text = element_text(size = 6), plot.title = element_text(size = 10), axis.text = element_text(size = 8))+ stat_compare_means(method = "t.test", comparisons = comparisons, paired = FALSE) + scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))
  plot_list_SDC1_bulk[[path]] <- p
  #print(p)
}# these stats/pvalues are NOT adjusted
plot_list_SDC1_bulk[["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1"]]
```

# SDC4 bulk
```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
comparisons <- list(c("SDC4_hi","SDC4_lo"))
plot_list_SDC4_bulk <- list()
for(i in seq_along(names(HALLMARK))){
  path <- paste0(names(HALLMARK[i]),"1")
  p <- VlnPlot(sobj, features = path, group.by = "SDC4_class_bulk", cols = class.cols, pt.size = 0) + geom_boxplot(width=0.1, fill="white") + NoLegend() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),axis.title.x=element_blank(),axis.title.y=element_blank(), legend.title = element_text(size = 8), legend.text = element_text(size = 6), plot.title = element_text(size = 10), axis.text = element_text(size = 8))+ stat_compare_means(method = "t.test", comparisons = comparisons, paired = FALSE) + scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))
  plot_list_SDC4_bulk[[path]] <- p
  #print(p)
}# these stats/pvalues are NOT adjusted
plot_list_SDC4_bulk[["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1"]]
```

# split response plots

# SDC1
```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
comparisons <- list(c("SDC1_hi","SDC1_lo"))
plot_list_SDC1_bulk <- list()
for(i in seq_along(names(HALLMARK))){
  path <- paste0(names(HALLMARK[i]),"1")
  p <- VlnPlot(sobj, features = path, group.by = "SDC1_class_bulk", split.by = "responseType", pt.size = 0, cols = response.cols) + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),axis.title.x=element_blank(),axis.title.y=element_blank(), legend.title = element_text(size = 8), legend.text = element_text(size = 6), plot.title = element_text(size = 10), axis.text = element_text(size = 8))
  plot_list_SDC1_bulk[[path]] <- p
  #print(p)
}# these stats/pvalues are NOT adjusted
plot_list_SDC1_bulk[["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1"]]
```

# SDC4
```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
comparisons <- list(c("SDC4_hi","SDC4_lo"))
plot_list_SDC4_bulk <- list()
for(i in seq_along(names(HALLMARK))){
  path <- paste0(names(HALLMARK[i]),"1")
  p <- VlnPlot(sobj, features = path, group.by = "SDC4_class_bulk", pt.size = 0, split.by = "responseType", cols = response.cols)+ theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),axis.title.x=element_blank(),axis.title.y=element_blank(), legend.title = element_text(size = 8), legend.text = element_text(size = 6), plot.title = element_text(size = 10), axis.text = element_text(size = 8))
  plot_list_SDC4_bulk[[path]] <- p
  #print(p)
}# these stats/pvalues are NOT adjusted
plot_list_SDC4_bulk[["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1"]]
```

# Plot correlation of SDC Expression vs EMT MS

# SDC1

# Responding

```{r warning=FALSE}
# subset object to just responding
sobj.responding <- subset(sobj, responseType == "responding")
meta <- sobj.responding@meta.data
```

# stat split plots
# SDC1
```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
comparisons <- list(c("SDC1_hi","SDC1_lo"))
plot_list_SDC1_R <- list()
for(i in seq_along(names(HALLMARK))){
  path <- paste0(names(HALLMARK[i]),"1")
  p <- VlnPlot(sobj.responding, features = path, group.by = "SDC1_class", pt.size = 0, cols = class.cols)+ geom_boxplot(width=0.1, fill="white") + NoLegend() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),axis.title.x=element_blank(),axis.title.y=element_blank(), legend.title = element_text(size = 8), legend.text = element_text(size = 6), plot.title = element_text(size = 10), axis.text = element_text(size = 8))+ stat_compare_means(method = "wilcox.test", comparisons = comparisons, paired = FALSE) + scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))
  plot_list_SDC1_R[[path]] <- p
  #print(p)
}# these stats/pvalues are NOT adjusted
plot_list_SDC1_R[["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1"]]
```

# SDC1 - bulk
```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
comparisons <- list(c("SDC1_hi","SDC1_lo"))
plot_list_SDC1_bulk_R <- list()
for(i in seq_along(names(HALLMARK))){
  path <- paste0(names(HALLMARK[i]),"1")
  p <- VlnPlot(sobj.responding, features = path, group.by = "SDC1_class_bulk", pt.size = 0, cols = class.cols)+ geom_boxplot(width=0.1, fill="white") + NoLegend() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),axis.title.x=element_blank(),axis.title.y=element_blank(), legend.title = element_text(size = 8), legend.text = element_text(size = 6), plot.title = element_text(size = 10), axis.text = element_text(size = 8))+ stat_compare_means(method = "wilcox.test", comparisons = comparisons, paired = FALSE) + scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))
  plot_list_SDC1_bulk_R[[path]] <- p
  #print(p)
}# these stats/pvalues are NOT adjusted
plot_list_SDC1_bulk_R[["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1"]]
```

# calculate spearman correlation
```{r}
cor_test <- cor.test(meta[["SDC1_exp"]], meta[['HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1']], method = "spearman") # or "pearson"
cor_test
```
# Plot correlation (with stats)
```{r fig.height=4, fig.width=6}
ggplot(meta, aes_string(x = 'SDC1_exp', y = 'HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1')) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "red") +
  labs(
    x = "SDC1 Expression",
    y = "EMT Module Score",
    title = paste0(
      "SDC1 vs EMT Module Score",
      " (r = ", round(cor_test$estimate, 2),
      ", p = ", signif(cor_test$p.value, 3), ") in Responding PDAC"
    )
  ) +
  theme_minimal() + theme(title = element_text(size = 8))
```

```{r}
VlnPlot(sobj, features = "SDC1", group.by = "responseType")
```

# Non-Responding

```{r warning=FALSE}
# subset object to just responding
sobj.NONresponding <- subset(sobj, responseType == "non-responding")
meta <- sobj.NONresponding@meta.data
```

# SDC1
```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
comparisons <- list(c("SDC1_hi","SDC1_lo"))
plot_list_SDC1_NR <- list()
for(i in seq_along(names(HALLMARK))){
  path <- paste0(names(HALLMARK[i]),"1")
  p <- VlnPlot(sobj.NONresponding, features = path, group.by = "SDC1_class", pt.size = 0, cols = class.cols)+ geom_boxplot(width=0.1, fill="white") + NoLegend() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),axis.title.x=element_blank(),axis.title.y=element_blank(), legend.title = element_text(size = 8), legend.text = element_text(size = 6), plot.title = element_text(size = 10), axis.text = element_text(size = 8))+ stat_compare_means(method = "wilcox.test", comparisons = comparisons, paired = FALSE) + scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))
  plot_list_SDC1_NR[[path]] <- p
  #print(p)
}# these stats/pvalues are NOT adjusted
plot_list_SDC1_NR[["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1"]]
```

# SDC1 - bulk
```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
comparisons <- list(c("SDC1_hi","SDC1_lo"))
plot_list_SDC1_bulk_NR <- list()
for(i in seq_along(names(HALLMARK))){
  path <- paste0(names(HALLMARK[i]),"1")
  p <- VlnPlot(sobj.NONresponding, features = path, group.by = "SDC1_class_bulk", pt.size = 0, cols = class.cols)+ geom_boxplot(width=0.1, fill="white") + NoLegend() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),axis.title.x=element_blank(),axis.title.y=element_blank(), legend.title = element_text(size = 8), legend.text = element_text(size = 6), plot.title = element_text(size = 10), axis.text = element_text(size = 8))+ stat_compare_means(method = "wilcox.test", comparisons = comparisons, paired = FALSE) + scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))
  plot_list_SDC1_bulk_NR[[path]] <- p
  #print(p)
}# these stats/pvalues are NOT adjusted
plot_list_SDC1_bulk_NR[["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1"]]
```

# calculate spearman correlation
```{r}
cor_test <- cor.test(meta[["SDC1_exp"]], meta[['HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1']], method = "spearman") # or "pearson"
cor_test
```
# Plot correlation (with stats)
```{r fig.height=4, fig.width=6}
ggplot(meta, aes_string(x = 'SDC1_exp', y = 'HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1')) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "red") +
  labs(
    x = "SDC1 Expression",
    y = "EMT Module Score",
    title = paste0(
      "SDC1 vs EMT Module Score",
      " (r = ", round(cor_test$estimate, 2),
      ", p = ", signif(cor_test$p.value, 3), ") in Non-Responding PDAC"
    )
  ) +
  theme_minimal() + theme(title = element_text(size = 8))
```

```{r}

```

# SDC4

# Responding

```{r warning=FALSE}
# subset object to just responding
sobj.responding <- subset(sobj, responseType == "responding")
meta <- sobj.responding@meta.data
```

# SDC4
```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
comparisons <- list(c("SDC4_hi","SDC4_lo"))
plot_list_SDC4_R <- list()
for(i in seq_along(names(HALLMARK))){
  path <- paste0(names(HALLMARK[i]),"1")
  p <- VlnPlot(sobj.responding, features = path, group.by = "SDC4_class", pt.size = 0, cols = class.cols)+ geom_boxplot(width=0.1, fill="white") + NoLegend() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),axis.title.x=element_blank(),axis.title.y=element_blank(), legend.title = element_text(size = 8), legend.text = element_text(size = 6), plot.title = element_text(size = 10), axis.text = element_text(size = 8))+ stat_compare_means(method = "wilcox.test", comparisons = comparisons, paired = FALSE) + scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))
  plot_list_SDC4_R[[path]] <- p
  #print(p)
}# these stats/pvalues are NOT adjusted
plot_list_SDC4_R[["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1"]]
```

# SDC4 - bulk
```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
comparisons <- list(c("SDC4_hi","SDC4_lo"))
plot_list_SDC4_bulk_R <- list()
for(i in seq_along(names(HALLMARK))){
  path <- paste0(names(HALLMARK[i]),"1")
  p <- VlnPlot(sobj.responding, features = path, group.by = "SDC4_class_bulk", pt.size = 0, cols = class.cols)+ geom_boxplot(width=0.1, fill="white") + NoLegend() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),axis.title.x=element_blank(),axis.title.y=element_blank(), legend.title = element_text(size = 8), legend.text = element_text(size = 6), plot.title = element_text(size = 10), axis.text = element_text(size = 8))+ stat_compare_means(method = "wilcox.test", comparisons = comparisons, paired = FALSE) + scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))
  plot_list_SDC4_bulk_R[[path]] <- p
  #print(p)
}# these stats/pvalues are NOT adjusted
plot_list_SDC4_bulk_R[["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1"]]
```

# calculate spearman correlation
```{r}
cor_test <- cor.test(meta[["SDC4_exp"]], meta[['HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1']], method = "spearman") # or "pearson"
cor_test
```
# Plot correlation (with stats)
```{r fig.height=4, fig.width=6}
ggplot(meta, aes_string(x = 'SDC4_exp', y = 'HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1')) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "red") +
  labs(
    x = "SDC4 Expression",
    y = "EMT Module Score",
    title = paste0(
      "SDC4 vs EMT Module Score",
      " (r = ", round(cor_test$estimate, 2),
      ", p = ", signif(cor_test$p.value, 3), ") in Responding PDAC"
    )
  ) +
  theme_minimal() + theme(title = element_text(size = 8))
```

```{r}
VlnPlot(sobj, features = "SDC4", group.by = "responseType")
```

# Non-Responding

```{r warning=FALSE}
# subset object to just responding
sobj.NONresponding <- subset(sobj, responseType == "non-responding")
meta <- sobj.NONresponding@meta.data
```

# SDC4
```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
comparisons <- list(c("SDC4_hi","SDC4_lo"))
plot_list_SDC4_NR <- list()
for(i in seq_along(names(HALLMARK))){
  path <- paste0(names(HALLMARK[i]),"1")
  p <- VlnPlot(sobj.NONresponding, features = path, group.by = "SDC4_class", pt.size = 0, cols = class.cols)+ geom_boxplot(width=0.1, fill="white") + NoLegend() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),axis.title.x=element_blank(),axis.title.y=element_blank(), legend.title = element_text(size = 8), legend.text = element_text(size = 6), plot.title = element_text(size = 10), axis.text = element_text(size = 8))+ stat_compare_means(method = "wilcox.test", comparisons = comparisons, paired = FALSE) + scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))
  plot_list_SDC4_NR[[path]] <- p
  #print(p)
}# these stats/pvalues are NOT adjusted
plot_list_SDC4_NR[["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1"]]
```

# SDC4 - bulk
```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
comparisons <- list(c("SDC4_hi","SDC4_lo"))
plot_list_SDC4_bulk_NR <- list()
for(i in seq_along(names(HALLMARK))){
  path <- paste0(names(HALLMARK[i]),"1")
  p <- VlnPlot(sobj.NONresponding, features = path, group.by = "SDC4_class_bulk", pt.size = 0, cols = class.cols)+ geom_boxplot(width=0.1, fill="white") + NoLegend() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),axis.title.x=element_blank(),axis.title.y=element_blank(), legend.title = element_text(size = 8), legend.text = element_text(size = 6), plot.title = element_text(size = 10), axis.text = element_text(size = 8))+ stat_compare_means(method = "wilcox.test", comparisons = comparisons, paired = FALSE) + scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))
  plot_list_SDC4_bulk_NR[[path]] <- p
  #print(p)
}# these stats/pvalues are NOT adjusted
plot_list_SDC4_bulk_NR[["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1"]]
```

# calculate spearman correlation
```{r}
cor_test <- cor.test(meta[["SDC4_exp"]], meta[['HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1']], method = "spearman") # or "pearson"
cor_test
```
# Plot correlation (with stats)
```{r fig.height=4, fig.width=6}
ggplot(meta, aes_string(x = 'SDC4_exp', y = 'HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION1')) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "red") +
  labs(
    x = "SDC4 Expression",
    y = "EMT Module Score",
    title = paste0(
      "SDC4 vs EMT Module Score",
      " (r = ", round(cor_test$estimate, 2),
      ", p = ", signif(cor_test$p.value, 3), ") in Non-Responding PDAC"
    )
  ) +
  theme_minimal() + theme(title = element_text(size = 8))
```

```{r}
sessionInfo()
```
